---
title: "PS-Solar_phase_2"
output: html_document
date: "2026-04-10"
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidycensus)
library(sf)
#library(tmap)
library(tmaptools)
library(DT)
library(ggplot2)
library(tidyverse)
library(janitor)
library(GGally)
library(DT)            # for visualizing data tables 
library(janitor)       # for automatically renaming variables
library(ggmap)         # for mapping points on maps
library(RColorBrewer)  # for color palettes
library(sf)            # for working with spatial data
library(leaflet)       # for highly customizable mapping
library(ggthemes)      # for more themes (including theme_map())
library(gplots)        # for col2hex() function
library(socviz)        # for the US county map
library(usmap)         # for plot_usmap
library(usdata)       # for abbr2state function
library(choroplethr)   # another option to make choropleths
library(choroplethrMaps)
library(maps)
library(rvest)
knitr::opts_chunk$set(echo = TRUE)
```

## Data loading and processing

```{r}
#imports Data
mass_solar_messy <- read_csv("C:/Users/pschnell/Downloads/Data Science sust/Solar-PV-Systems-in-MA-as-of-Jan-2025-readable.csv", 
    col_types = cols(`Date In Service` = col_date(format = "%m/%d/%Y"), 
        `Total Cost with Design Fees` = col_number(), 
        `Total Grant` = col_number()))

#cleans names
mass_solar_clean<-mass_solar_messy |>
  janitor::clean_names()|>
   mutate(year=year(date_in_service), month=month(date_in_service),weekday=wday(date_in_service), day=day(date_in_service))


```

Income data from WIki
```{r}
#income data
url<-"https://en.wikipedia.org/wiki/List_of_Massachusetts_locations_by_per_capita_income"
webpage<-rvest::read_html(url)

county <- webpage |>
  html_nodes("td:nth-child(3)") |>
  html_text2() 

median_income <- webpage |>
  html_nodes("td:nth-child(4)") |>
  html_text2() 

county_income <- tibble(county=county,median_income=median_income)


url<-"https://en.wikipedia.org/wiki/List_of_Massachusetts_locations_by_per_capita_income"
webpage<-rvest::read_html(url)

town <- webpage |>
  html_nodes("td:nth-child(2)") |>
  html_text2() 


median_income_town <- webpage |>
  html_nodes("td:nth-child(5)") |>
  html_text2() 
  
  

```

citys/towns
```{r}
#income data
url<-"https://en.wikipedia.org/wiki/List_of_municipalities_in_Massachusetts"
webpage<-rvest::read_html(url)

city <- webpage |>
  html_nodes(".static-row-numbers td:nth-child(1)") |>
  html_text2() 

county <- webpage |>
  html_nodes(".static-row-numbers td:nth-child(3)") |>
  html_text2() 

type <- webpage |>
  html_nodes(".static-row-numbers td:nth-child(2)") |>
  html_text2() 

pop <- webpage |>
  html_nodes(".static-row-numbers td:nth-child(5)") |>
  html_text2() 

land_area <- webpage |>
  html_nodes(".static-row-numbers td:nth-child(7)") |>
  html_text2() 

mass_city <- tibble(county=county,city=city, pop_2020=pop, type=type, land_area=land_area)

mass_city

mass_city<-mass_city|>
 # grepl("seat", county)
  mutate(county=sub("(seat)"," ", county))|>
  mutate(county=sub("\\("," ", county))|>
  mutate(county=sub("\\)"," ", county))



mass_city_1<-mass_solar_clean|>
  select("city", "county")|>
  unique()
  
  mass_city
  
  na<- mass_city|>
    filter(is.na(county))
  
```

```{r}
# install.packages(c("ggplot2", "dplyr"))
library(ggplot2)
library(dplyr)

# Timeline data
solar_programs <- data.frame(
  program = c(
    "Federal Solar Tax Credit (ITC)",
    "MA Solar Tax Credit",
    "MassCEC Commonwealth Solar",
    "SREC I",
    "SREC II",
    "Mass Solar Loan",
    "SMART (Mass SMART)",
    "EmPower Grants"
  ),
  
  start = as.Date(c(
    "2006-01-01",
    "1979-01-01",
    "2008-01-01",
    "2010-01-01",
    "2014-04-25",
    "2015-06-01",
    "2018-11-26",
    "2020-01-01"
  )),
  
  end = as.Date(c(
    "2034-12-31",
    "2030-12-31",
    "2015-12-31",
    "2014-04-24",
    "2018-11-26",
    "2020-12-31",
    "2030-12-31",
    "2030-12-31"
  ))
)

# Order programs top-to-bottom
solar_programs$program <- factor(
  solar_programs$program,
  levels = rev(solar_programs$program)
)

# Plot
ggplot(solar_programs) +
  geom_segment(
    aes(
      x = start,
      xend = end,
      y = program,
      yend = program
    ),
    linewidth = 8,
    color = "steelblue"
  ) +
  labs(
    title = "Massachusetts & Federal Solar Incentive Timeline",
    x = "Year",
    y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.y = element_text(face = "bold")
  )
```
```{r}
# Timeline data
solar_programs <- data.frame(
  program = c(
    "Federal Solar Tax Credit (ITC)",
    "MA Solar Tax Credit",
    "MassCEC Commonwealth Solar",
    "SREC I",
    "SREC II",
    "Mass Solar Loan",
    "SMART (Mass SMART)",
    "EmPower Grants"
  ),
  
  start = as.Date(c(
    "2006-01-01",
    "1979-01-01",
    "2008-01-01",
    "2010-01-01",
    "2014-04-25",
    "2015-06-01",
    "2018-11-26",
    "2020-01-01"
  )),
  
  end = as.Date(c(
    "2034-12-31",
    "2030-12-31",
    "2015-12-31",
    "2014-04-24",
    "2018-11-26",
    "2020-12-31",
    "2030-12-31",
    "2030-12-31"
  ))
)

# Order programs top-to-bottom
solar_programs$program <- factor(
  solar_programs$program,
  levels = rev(solar_programs$program)
)

# Plot
ggplot(solar_programs) +
  geom_segment(
    aes(
      x = start,
      xend = end,
      y = program,
      yend = program
    ),
    linewidth = 8,
    color = "blue"
  ) +
  labs(
    title = "Massachusetts & Federal Solar Incentive Timeline",
    x = "Year",
    y = NULL
  ) +
  theme_clean(base_size = 13) +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.y = element_text(face = "bold")
  )
```
```{r}
# install.packages(c("ggplot2", "dplyr", "scales"))
library(ggplot2)
library(dplyr)
library(scales)

# Timeline data
solar_programs <- data.frame(
  program = c(
    "Federal Solar Tax Credit (Carter Era)",
    "Federal Solar Tax Credit (ITC)",
    "MA Solar Tax Credit",
    "MassCEC Commonwealth Solar",
    "SREC I",
    "SREC II",
    "Mass Solar Loan",
    "SMART (Mass SMART)",
    "EmPower Grants"
  ),

  start = as.Date(c(
    "1978-01-01",
    "2006-01-01",
    "1979-01-01",
    "2008-01-01",
    "2010-01-01",
    "2014-04-25",
    "2015-06-01",
    "2018-11-26",
    "2020-01-01"
  )),

  end = as.Date(c(
    "1985-12-31",
    "2034-12-31",
    "2030-12-31",
    "2015-12-31",
    "2014-04-24",
    "2018-11-26",
    "2020-12-31",
    "2030-12-31",
    "2030-12-31"
  ))
)

# Labels and midpoints
solar_programs <- solar_programs %>%
  mutate(
    label = paste0(
      format(start, "%Y"),
      "–",
      format(end, "%Y")
    ),
    mid = start + (end - start) / 2
  )

# Reverse order for plotting
solar_programs$program <- factor(
  solar_programs$program,
  levels = rev(solar_programs$program)
)

# Plot
ggplot(solar_programs) +

  # Narrow bars
  geom_segment(
    aes(
      x = start,
      xend = end,
      y = program,
      yend = program
    ),
    linewidth = 4,
    color = "steelblue"
  ) +

  # Labels inside bars
  geom_text(
    aes(
      x = mid,
      y = program,
      label = label
    ),
    color = "white",
    fontface = "bold",
    size = 3.3
  ) +

  # Yearly x-axis labels and grid
  scale_x_date(
    date_breaks = "2 year",
    date_minor_breaks = "2 year",
    labels = label_date("%Y"),
    expand = expansion(mult = c(0.01, 0.01))
  ) +

  labs(
    title = "Massachusetts & Federal Solar Incentive Timeline",
    subtitle = "Major Massachusetts and Federal Solar Incentive Programs",
    x = "Year",
    y = NULL
  ) +

  theme_minimal(base_size = 12) +

  theme(
    panel.grid.major.x = element_line(color = "grey75", linewidth = 0.45),
    panel.grid.minor.x = element_line(color = "grey90", linewidth = 0.25),
    panel.grid.major.y = element_line(color = "grey92"),
    panel.grid.minor.y = element_blank(),

    axis.text.x = element_text(
      angle = 90,
      vjust = 0.5,
      size = 8
    ),

    axis.text.y = element_text(face = "bold"),

    plot.title = element_text(
      face = "bold",
      size = 16
    ),

    plot.subtitle = element_text(size = 11)
  )
```
```{r}
# install.packages(c("ggplot2", "dplyr", "scales"))
library(ggplot2)
library(dplyr)
library(scales)

# Timeline data
solar_programs <- data.frame(
  program = c(
    "Federal Solar Tax Credit (“Energy Tax Act)",
    "Federal Solar Tax Credit (ITC)",
    "MA Solar Tax Credit",
    "MassCEC Commonwealth Solar",
    "SREC I",
    "SREC II",
    "Mass Solar Loan",
    "SMART (Mass SMART)",
    "EmPower Grants"
  ),

  start = as.Date(c(
    "1978-01-01",
    "2006-01-01",
    "1979-01-01",
    "2008-01-01",
    "2010-01-01",
    "2014-04-25",
    "2015-06-01",
    "2018-11-26",
    "2020-01-01"
  )),

  end = as.Date(c(
    "1985-12-31",
    "2034-12-31",
    "2030-12-31",
    "2015-12-31",
    "2014-04-24",
    "2018-11-26",
    "2020-12-31",
    "2030-12-31",
    "2030-12-31"
  ))
)

# Reverse order for plotting
solar_programs$program <- factor(
  solar_programs$program,
  levels = rev(solar_programs$program)
)

# Every-other-year breaks starting at 1978
year_breaks <- seq(
  as.Date("1978-01-01"),
  as.Date("2034-01-01"),
  by = "2 years"
)

# Plot
ggplot(solar_programs) +

  # Narrow timeline bars
  geom_segment(
    aes(
      x = start,
      xend = end,
      y = program,
      yend = program
    ),
    linewidth = 4,
    color = "steelblue"
  ) +

  # X-axis formatting
  scale_x_date(
    breaks = year_breaks,
    date_minor_breaks = "1 year",
    labels = label_date("%Y"),
    expand = expansion(mult = c(0.01, 0.01))
  ) +

  labs(
    title = "Massachusetts & Federal Solar Incentive Timeline",
    x = "Year",
    y = NULL
  ) +

  theme_minimal(base_size = 12) +

  theme(
    panel.grid.major.x = element_line(
      color = "grey75",
      linewidth = 0.45
    ),

    panel.grid.minor.x = element_line(
      color = "grey90",
      linewidth = 0.25
    ),

    panel.grid.major.y = element_line(color = "grey92"),
    panel.grid.minor.y = element_blank(),

    axis.text.x = element_text(
      angle = 90,
      vjust = 0.5,
      size = 8
    ),

    axis.text.y = element_text(
      face = "bold",
      size = 10
    ),

    plot.title = element_text(
      face = "bold",
      size = 16
    ),

    plot.subtitle = element_text(size = 11)
  )
```

```{r}
# install.packages("ggplot2")

library(ggplot2)

# SMART incentive values
values <- c(
  0.31126, 0.29881, 0.28686, 0.27538,
  0.26437, 0.25379, 0.24364, 0.23390,
  0.22454, 0.21556, 0.20694, 0.19866,
  0.19071, 0.18308, 0.17576, 0.16873
)

# Create dataframe
smart_df <- data.frame(
  Block = 1:16,
  Incentive = values
)

# Plot
ggplot(smart_df, aes(x = Block, y = Incentive)) +

  geom_line(
    color = "steelblue",
    linewidth = 1.2
  ) +

  geom_point(
    color = "darkblue",
    size = 2.5
  ) +

  geom_text(
    aes(label = round(Incentive, 3)),
    vjust = -0.8,
    size = 3
  ) +

  scale_x_continuous(
    breaks = 1:16
  ) +

  labs(
    title = "Mass SMART Incentive Decline by Block",
    subtitle = "Allocated for Mass Elec National Grid ≤ 25 kW AC — 200% Compensation Rate — 10-Year Term",
    x = "Capacity Block (every 18kw",
    y = "Incentive Rate ($/kWh)"
  ) +

  theme_minimal(base_size = 13) +

  theme(
    panel.grid.minor = element_line(color = "grey92"),
    panel.grid.major = element_line(color = "grey80"),

    plot.title = element_text(
      face = "bold",
      size = 16
    ),

    axis.text = element_text(size = 10)
  )
```
```{r}
# install.packages(c("ggplot2", "dplyr"))

library(ggplot2)
library(dplyr)

# SMART incentive values ($/kWh)
incentive_values <- c(
  0.31126, 0.29881, 0.28686, 0.27538,
  0.26437, 0.25379, 0.24364, 0.23390,
  0.22454, 0.21556, 0.20694, 0.19866,
  0.19071, 0.18308, 0.17576, 0.16873
)

# Capacity block sizes converted to cumulative MW
capacity_values <- c(
  18.004, 18.004, 18.004, 18.004,
  18.004, 18.004, 18.004, 18.004,
  18.093, 18.093, 18.093, 18.093,
  18.093, 18.093, 18.093, 18.093
)

# Build dataframe
smart_df <- data.frame(
  Block = 1:16,
  Incentive = incentive_values,
  Capacity_MW = cumsum(capacity_values)
)

# Plot
ggplot(smart_df, aes(x = Capacity_MW, y = Incentive)) +

  # Line
  geom_line(
    color = "steelblue",
    linewidth = 1.3
  ) +

  # Points
  geom_point(
    color = "darkblue",
    size = 3
  ) +

  # Incentive labels
  geom_text(
    aes(label = round(Incentive, 3)),
    vjust = -1,
    size = 3
  ) +

  scale_x_continuous(
    breaks = seq(0, 300, by = 25),
    minor_breaks = seq(0, 300, by = 12.5)
  ) +

  labs(
    title = "Mass SMART Incentive Decline vs Installed Capacity",
    subtitle = "Massachusetts Electric (National Grid)\n≤ 25 kW AC — 200% Compensation Rate — 10-Year Term",
    x = "Cumulative Capacity Installed (MW)",
    y = "Incentive Rate ($/kWh)"
  ) +

  theme_minimal(base_size = 13) +

  theme(
    panel.grid.major = element_line(
      color = "grey80",
      linewidth = 0.5
    ),

    panel.grid.minor = element_line(
      color = "grey92",
      linewidth = 0.3
    ),

    plot.title = element_text(
      face = "bold",
      size = 16
    ),

    plot.subtitle = element_text(size = 11),

    axis.text = element_text(size = 10)
  )

```

```{r}
# install.packages(c("ggplot2", "dplyr"))

library(ggplot2)
library(dplyr)

# SMART incentive values converted from $/kWh to $/MWh
# (multiply by 1000)
incentive_values <- c(
  0.31126, 0.29881, 0.28686, 0.27538,
  0.26437, 0.25379, 0.24364, 0.23390,
  0.22454, 0.21556, 0.20694, 0.19866,
  0.19071, 0.18308, 0.17576, 0.16873
) * 1000

# Capacity block sizes (MW)
capacity_values <- c(
  18.004, 18.004, 18.004, 18.004,
  18.004, 18.004, 18.004, 18.004,
  18.093, 18.093, 18.093, 18.093,
  18.093, 18.093, 18.093, 18.093
)

# Build dataframe
smart_df <- data.frame(
  Block = 1:16,
  Incentive_MWh = incentive_values,
  Capacity_MW = cumsum(capacity_values)
)

# Plot
ggplot(smart_df, aes(x = Capacity_MW, y = Incentive_MWh)) +

  # Line
  geom_line(
    color = "steelblue",
    linewidth = 1.3
  ) +

  # Points
  geom_point(
    color = "darkblue",
    size = 3
  ) +

  # Labels
  geom_text(
    aes(label = round(Incentive_MWh, 1)),
    vjust = -1,
    size = 3
  ) +

  scale_x_continuous(
    breaks = seq(0, 300, by = 25),
    minor_breaks = seq(0, 300, by = 12.5)
  ) +

  scale_y_continuous(
    breaks = seq(150, 325, by = 25),
    minor_breaks = seq(150, 325, by = 12.5)
  ) +

  labs(
    title = "Mass SMART Incentive Decline vs Installed Capacity",
    subtitle = "Massachusetts Electric (National Grid)n/ ≤ 25 kW AC — 200% Compensation Rate — 10-Year Term",
    x = "Cumulative Capacity Installed (MW)",
    y = "Incentive Rate ($/MWh)"
  ) +

  theme_minimal(base_size = 13) +

  theme(
    panel.grid.major = element_line(
      color = "grey80",
      linewidth = 0.5
    ),

    panel.grid.minor = element_line(
      color = "grey92",
      linewidth = 0.3
    ),

    plot.title = element_text(
      face = "bold",
      size = 16
    ),

    plot.subtitle = element_text(size = 11),

    axis.text = element_text(size = 10)
  )
```
```{r}
# install.packages(c("ggplot2", "dplyr"))

library(ggplot2)
library(dplyr)

# Incentive values converted to $/MWh
incentive_values <- c(
  0.31126, 0.29881, 0.28686, 0.27538,
  0.26437, 0.25379, 0.24364, 0.23390,
  0.22454, 0.21556, 0.20694, 0.19866,
  0.19071, 0.18308, 0.17576, 0.16873
) * 1000

# Capacity (MW)
capacity_values <- c(
  18.004, 18.004, 18.004, 18.004,
  18.004, 18.004, 18.004, 18.004,
  18.093, 18.093, 18.093, 18.093,
  18.093, 18.093, 18.093, 18.093
)

# Data frame
smart_df <- data.frame(
  Block = 1:16,
  Incentive_MWh = incentive_values,
  Capacity_MW = capacity_values
)

# Colors
line_blue <- "lightblue"
point_blue <- "darkblue"
grey_text <- "black"

ggplot(smart_df, aes(x = Block, y = Incentive_MWh)) +

  # Line
  geom_line(
    color = line_blue,
    linewidth = 1.3
  ) +

  # Points
  geom_point(
    color = point_blue,
    size = 3
  ) +

  # Labels shifted right + down
  geom_text(
    aes(label = round(Incentive_MWh, 1)),
    nudge_x = 0.25,
    nudge_y = -5,
    size = 3,
    color = grey_text
  ) +

  scale_x_continuous(
    breaks = 1:16
  ) +

  scale_y_continuous(
    breaks = seq(150, 325, by = 25),
    minor_breaks = seq(150, 325, by = 12.5)
  ) +

  labs(
    title = "Mass SMART Incentive Decline by Block",
    subtitle = "Massachusetts Electric (National Grid)\n≤ 25 kW AC — 200% Compensation Rate — 10-Year Term",
    x = "Block",
    y = "Incentive Rate ($/MWh)"
  ) +

  theme_clean(base_size = 13) +

  theme(
    panel.grid.major = element_line(color = "grey85"),
    panel.grid.minor = element_line(color = "grey92"),

    plot.title = element_text(
      face = "bold",
      size = 16
    ),

    plot.subtitle = element_text(size = 11),

    axis.text = element_text(size = 10)
  )

```


```{r}
# install.packages(c("ggplot2", "dplyr"))

library(ggplot2)
library(dplyr)

# SMART incentive values
incentive_values <- c(
  0.31126, 0.29881, 0.28686, 0.27538,
  0.26437, 0.25379, 0.24364, 0.23390,
  0.22454, 0.21556, 0.20694, 0.19866,
  0.19071, 0.18308, 0.17576, 0.16873
)

# Capacity block sizes (MW)
capacity_values <- c(
  18.004, 18.004, 18.004, 18.004,
  18.004, 18.004, 18.004, 18.004,
  18.093, 18.093, 18.093, 18.093,
  18.093, 18.093, 18.093, 18.093
)

# Create dataframe
smart_df <- data.frame(
  Block = 1:16,
  Incentive = incentive_values,
  Capacity_MW = capacity_values
)

# Plot
ggplot(smart_df, aes(x = Block, y = Incentive)) +

  # Line
  geom_line(
    color = "steelblue",
    linewidth = 1.2
  ) +

  # Points scaled by capacity
  geom_point(
    aes(size = Capacity_MW),
    color = "darkblue",
    alpha = 0.9
  ) +

  # Incentive labels
  geom_text(
    aes(label = round(Incentive, 3)),
    vjust = -1.0,
    size = 3
  ) +

  # Capacity labels
  geom_text(
    aes(label = paste0(Capacity_MW, " MW")),
    vjust = 1.8,
    color = "darkred",
    size = 3
  ) +

  scale_x_continuous(
    breaks = 1:16
  ) +

  scale_size_continuous(
    range = c(4, 7),
    name = "Capacity Block Size (MW)"
  ) +

  labs(
    title = "Mass SMART Incentive Decline by Capacity Block",
    subtitle = "Massachusetts Electric (National Grid)\n≤ 25 kW AC — 200% Compensation Rate — 10-Year Term",
    x = "Capacity Block",
    y = "Incentive Rate ($/kWh)"
  ) +

  theme_minimal(base_size = 13) +

  theme(
    panel.grid.minor = element_line(color = "grey92"),
    panel.grid.major = element_line(color = "grey80"),

    plot.title = element_text(
      face = "bold",
      size = 16
    ),

    plot.subtitle = element_text(size = 11),

    axis.text = element_text(size = 10)
  )
```


## % Grants

```{r}
mass_solar_clean
mass_solar_clean_prec<- mass_solar_clean|>
  filter(total_cost_with_design_fees!=0)|>
  mutate(grant_prec = 100*(total_grant/total_cost_with_design_fees))|>
  arrange(desc(grant_prec))|>
  select(-"installer", -"module_manufacturer", -"inverter_manufacturer", -"meter_manufacturer", -"utility", -"x3rd_party_owner", -"srec_eligible", -"estimated_annual_production_k_whr")

mass_solar_clean_prec<-mass_solar_clean_prec[-c(1,2,3),]
  
 # filter(grant_prec >= 0)
#    filter(!is.na(as.numeric(grant_prec))\

 #mass_solar_clean_prec$grant_prec <- gsub("\\D", "", mass_solar_clean_prec$grant_prec)

 mass_solar_clean_prec$grant_prec<- str_replace_all(mass_solar_clean_prec$grant_prec,"NaN","0")
  mass_solar_clean_prec$grant_prec<- str_replace_all(mass_solar_clean_prec$grant_prec,"NA","0")
 
   which(is.na(as.character( mass_solar_clean_prec$grant_prec)))

  

mass_solar_clean_prec

```

```{r}
#Counting solar
solar_count_num<-mass_solar_clean|>
  group_by(county)|>
  count(facility_type)|>
 arrange(desc(n))

solar_count_city<-mass_solar_clean|>
  group_by(city)|>
  count(facility_type)|>
 arrange(desc(n))

solar_count_city

solar_count_num_resid<-solar_count_num|>
  filter(facility_type=="Residential (3 or fewer dwelling units per building)")


solar_count_resid_city<-solar_count_city|>
  filter(facility_type=="Residential (3 or fewer dwelling units per building)")

solar_count_resid4_city<-solar_count_city|>
  filter(facility_type=="Residential (3 or fewer dwelling units per building)")


```
```{r}

#filters for single family
 just_residential<-mass_solar_clean_prec|>
 filter(facility_type=="Residential (3 or fewer dwelling units per building)") 
 just_residential
 
 
 #creates 
mean_prec_grant_single_fam<-just_residential |>
  group_by(county) |>
  summarise(mean_prec_grant=mean(as.numeric(grant_prec))) |>
   arrange(desc(mean_prec_grant)) 

  mean_prec_grant_single_fam
  
 solar_resid_prec_grant<- solar_count_num_resid |>
   left_join(mean_prec_grant_single_fam)
  
 solar_resid_prec_grant
 
 
ggplot(solar_resid_prec_grant, aes(x = n, y=mean_prec_grant, label=county))+
 geom_point()+
  ggtitle("Comparison of average grant precentage recieved and number of single family residential solar per county")+
labs(y="Precentage covered with grant", x="Number of Single Residence Solar Permits")+
  geom_text(nudge_y = 0.2)
```
Repeats for multi family:
```{r}
just_residential4_prec<-mass_solar_clean_prec|>
 filter(facility_type=="Multi-family residential (4 or more dwelling units per building)") 
 
 #creates 
mean_grant_prec_multi_fam<-just_residential4_prec |>
  group_by(county) |>
  summarise(mean_grant_prec=mean(as.numeric(grant_prec))) |>
   arrange(desc(mean_grant_prec)) 
  
 solar_resid4_grant<- solar_count_num_resid |>
   left_join(mean_grant_prec_multi_fam)
  

 
ggplot(solar_resid4_grant, aes(x = n, y=mean_grant_prec, label=county))+
 geom_point()+
  ggtitle("Comparison of average grant recieved and number of multi family residential solar per county")+
labs(y="Average presentage covered with grant", x="Number of Multi Residence Solar Permits")+
  geom_text(nudge_y = 0.9)
```
Single Familt facets:
```{r}

just_res4_town<-mass_solar_clean_prec|>
 filter(facility_type=="Multi-family residential (4 or more dwelling units per building)") 
 
 #creates 
mean_grant4_city<-just_res4_town |>
  group_by(city) |>
  summarise(mean_grant_prec=mean(as.numeric(grant_prec))) |>
   arrange(desc(mean_grant_prec)) |>
  filter(mean_grant_prec!=0)

  
mean_grant4_city

 solar_resid4_city_grant<- solar_count_resid4_city |>
   inner_join(mean_grant4_city)
  
  
solar_resid4_city_grant<- solar_resid4_city_grant|>
inner_join(mass_city)
  
solar_resid4_city_grant<- solar_resid4_city_grant|>
  mutate(county1=sub("\\s+$", "", county))

#unique(solar_resid4_city_grant)
 
ggplot(solar_resid4_city_grant, aes(x = n, y=mean_grant_prec, label=city))+
 geom_point()+
  facet_grid(~county1)+
  ggtitle("Comparison of average grant recieved and number of multi family residential solar per county")+
labs(y="Average presentage covered with grant", x="Number of Multi Residence Solar Permits")+
  geom_text(nudge_x = 3000, nudge_y = 3)
  


```


Multi Family Facets
```{r}

just_res4_town<-mass_solar_clean_prec|>
filter(facility_type=="Multi-family residential (4 or more dwelling units per building)") 
 
 #creates 
mean_grant4_city<-just_res4_town |>
  group_by(city) |>
  summarise(mean_grant_prec=mean(as.numeric(grant_prec))) |>
   arrange(desc(mean_grant_prec)) |>
  filter(mean_grant_prec!=0)

  
mean_grant4_city

 solar_resid4_city_grant<- solar_count_resid4_city |>
   inner_join(mean_grant4_city)
  
  
solar_resid4_city_grant<- solar_resid4_city_grant|>
inner_join(mass_city)
  
solar_resid4_city_grant<- solar_resid4_city_grant|>
  mutate(county1=sub("\\s+$", "", county))

#unique(solar_resid4_city_grant)
 
ggplot(solar_resid4_city_grant, aes(x = n, y=mean_grant_prec, label=city))+
 geom_point()+
  facet_grid(~county1)+
  ggtitle("Comparison of average grant recieved and number of multi family residential solar per county")+
labs(y="Average presentage covered with grant", x="Number of Multi Residence Solar Permits")+
  geom_text(nudge_x = 3000, nudge_y = 3)
  


```

Focus on Provincetown:


```{r}
just_ptown<- mass_solar_clean|>
  filter(city=="Provincetown")|>
  filter(facility_type=="Residential (3 or fewer dwelling units per building)") |>
  mutate(has_grant= total_grant>0)|>
  arrange(desc(total_grant))

just_ptown

just_ptown_year<-just_ptown|>
  count(year)|>
 mutate(num_instal=n)|>
  select(-"n")

grants_ptown<-just_ptown|>
  group_by(year)|>
  summarise(count=sum(has_grant, na.rm=TRUE))
  
ptown_grant_year<-just_ptown_year|>
  left_join(grants_ptown)

ptown_grant_year
  
ggplot(ptown_grant_year, aes(x=year))+
  geom_col(aes(y=num_instal))+
  geom_col(aes(y=count), fill="green")+
  ggtitle("Provincetown single family solar instalations over time, with grant recieved ")+
labs(y="Number of Instalations", x="Year", fill="Number covered by grants")


```
```{r}
just_ptown<- mass_solar_clean|>
  filter(city=="Provincetown")|>
  filter(facility_type=="Multi-family residential (4 or more dwelling units per building)") |>
  mutate(has_grant= total_grant>0)|>
  arrange(desc(total_grant))

just_ptown

just_ptown_year<-just_ptown|>
  count(year)|>
 mutate(num_instal=n)|>
  select(-"n")

grants_ptown<-just_ptown|>
  group_by(year)|>
  summarise(count=sum(has_grant, na.rm=TRUE))
  
ptown_grant_year<-just_ptown_year|>
  left_join(grants_ptown)

ptown_grant_year
  
ggplot(ptown_grant_year, aes(x=year))+
  geom_col(aes(y=num_instal))+
  geom_col(aes(y=count), fill="green")+
  ggtitle("Provincetown multi family solar instalations over time, with grant recieved ")+
labs(y="Number of Instalations", x="Year", fill="Number covered by grants")
```



- group by multiple catagories,, and then use count-- make grid of all the graphs 
```{r}
all_town<-mass_solar_clean|>
  mutate(has_grant= total_grant>0)|>
  group_by(city, year)|>
  summarise(count=n(), grant=sum(has_grant, na.rm=TRUE))




all_town<-all_town|>
  mutate(year=as.numeric(year))
all_town

library(plotly)
p <- ggplot(all_town, aes(x=year))+
  geom_col(aes(y=count))+
  geom_col(aes(y=grant), fill="green")+
  facet_wrap(~city)+
  ggtitle("Provincetown multi family solar instalations over time, with grant recieved ")+
labs(y="Number of Instalations", x="Year", fill="Number covered by grants")
  

ggplotly(p)



```
```{r}
total_solar<-mass_solar_clean|>
  mutate(has_grant= total_grant>0)|>
  group_by(year)|>
  summarise(count=n(), grant=sum(has_grant, na.rm=TRUE))

total_solar

  library(plotly)

p <- ggplot(total_solar, aes(x=year))+
  geom_col(aes(y=count))+
  geom_col(aes(y=grant), fill="green")+
  ggtitle("Mass total solar instalations over time, with grant recieved ")+
labs(y="Number of Instalations", x="Year", fill="Number covered by grants")
  

ggplotly(p)


```

```{r} 
mass_solar_clean

total_solar_type<-mass_solar_clean|>
  mutate(has_grant= total_grant>0)|>
  group_by(facility_type, year)|>
  summarise(count=n(), grant=sum(has_grant, na.rm=TRUE))

total_solar_type

  library(plotly)

p <- ggplot(total_solar_type, aes(x=year))+
  geom_col(aes(y=count))+
  geom_col(aes(y=grant), fill="green")+
  facet_wrap(~facility_type, scales="free")
  ggtitle("Mass total solar instalations over time, with grant recieved ")+
labs(y="Number of Instalations", x="Year", fill="Number covered by grants")
  

ggplotly(p)

```

- geom text.. xy coordinates make notes of fedral events and things to lable... 
chatty crap

```{r}

# ================================

# Massachusetts Solar + SREC Analysis

# ================================

library(dplyr)
library(ggplot2)
library(lubridate)
library(tibble)

# -------------------------------

# 1. SREC price dataset (DOER-based)

# -------------------------------

ma_srec <- tibble(
year = 2010:2025,

# SREC-I ACP (ceiling)

acp_srec1 = c(
600, 550, 550, 550,
523, 496, 472, 448,
426, 404, 384, 365,
347, 330, 330, 330
),

# SREC-II ACP (NA before 2014)

acp_srec2 = c(
NA, NA, NA, NA,
375, 375, 350, 350,
350, 333, 316, 300,
285, 271, 257, 244
),

# Auction clearinghouse price (floor before fee)

auction_price = c(
300, 300, 300, 300,
300, 300, 300,
285, 271, 257,
244, 232, 221,
210, 199, 189
),

# Net to seller (after 5% fee)

auction_net = auction_price * 0.95
)

# -------------------------------

# 2. Prepare installations data

# -------------------------------

installs_yearly <- mass_solar_clean |>
filter(
facility_type == "Residential (3 or fewer dwelling units per building)",
year >= 2010
) |>
mutate(year = year(as.Date(date_in_service))) |>
count(year, name = "num_instal")

# -------------------------------

# 3. Merge datasets

# -------------------------------

plot_data <- installs_yearly |>
left_join(ma_srec, by = "year")

# -------------------------------

# 4. Scale factor for dual axis

# -------------------------------

scale_factor <- max(plot_data$num_instal, na.rm = TRUE) /
max(plot_data$auction_net, na.rm = TRUE)

# -------------------------------

# 5. Visualization

# -------------------------------

ggplot(plot_data, aes(x = year)) +

# Installations (bars)

geom_col(aes(y = num_instal),
fill = "steelblue", alpha = 0.8) +

# SREC price band (floor → ceiling)

geom_ribbon(
aes(
ymin = auction_net * scale_factor,
ymax = acp_srec2 * scale_factor
),
fill = "orange",
alpha = 0.3
) +

# Floor price (actual payout)

geom_line(
aes(y = auction_net * scale_factor),
color = "black",
linewidth = 1.2
) +

geom_point(
aes(y = auction_net * scale_factor),
color = "black"
) +

# Dual axis

scale_y_continuous(
name = "Number of Installations",
sec.axis = sec_axis(~ . / scale_factor, name = "SREC Price ($)")
) +

scale_x_continuous(breaks = seq(2010, 2025, 1)) +

labs(
title = "Massachusetts Solar Installations vs SREC Price Range",
subtitle = "Bars = installations | Black line = SREC auction value | Orange band = price range",
x = "Year"
) +

theme_minimal()


```
```{r}

# ================================

# Massachusetts Solar + SREC Analysis

# ================================

library(dplyr)
library(ggplot2)
library(lubridate)
library(tibble)

# -------------------------------

# 1. SREC price dataset (DOER-based)

# -------------------------------

ma_srec <- tibble(
year = 2010:2025,

# SREC-I ACP (ceiling)

acp_srec1 = c(
600, 550, 550, 550,
523, 496, 472, 448,
426, 404, 384, 365,
347, 330, 330, 330
),

# SREC-II ACP (NA before 2014)

acp_srec2 = c(
NA, NA, NA, NA,
375, 375, 350, 350,
350, 333, 316, 300,
285, 271, 257, 244
),

# Auction clearinghouse price (floor before fee)

auction_price = c(
300, 300, 300, 300,
300, 300, 300,
285, 271, 257,
244, 232, 221,
210, 199, 189
),

# Net to seller (after 5% fee)

auction_net = auction_price * 0.95
)

# -------------------------------

# 2. Prepare installations data

# -------------------------------

installs_yearly <- mass_solar_clean |>
filter(
facility_type == "Residential (3 or fewer dwelling units per building)",
year >= 2010
) |>
mutate(year = year(as.Date(date_in_service))) |>
count(year, name = "num_instal")

# -------------------------------

# 3. Merge datasets

# -------------------------------

plot_data <- installs_yearly |>
left_join(ma_srec, by = "year")

# -------------------------------

# 4. Scale factor for dual axis

# -------------------------------

scale_factor <- max(plot_data$num_instal, na.rm = TRUE) /
max(plot_data$auction_net, na.rm = TRUE)

# -------------------------------

# 5. Visualization

# -------------------------------

ggplot(plot_data, aes(x = year)) +

# Installations (bars)

geom_col(aes(y = num_instal),
fill = "steelblue", alpha = 0.8) +

# SREC price band (floor → ceiling)

geom_ribbon(
aes(
ymin = auction_net * scale_factor,
ymax = acp_srec2 * scale_factor
),
fill = "orange",
alpha = 0.3
) +

# Floor price (actual payout)

geom_line(
aes(y = auction_net * scale_factor),
color = "black",
linewidth = 1.2
) +

geom_point(
aes(y = auction_net * scale_factor),
color = "black"
) +

# Dual axis

scale_y_continuous(
name = "Number of Installations",
sec.axis = sec_axis(~ . / scale_factor, name = "SREC Price ($)")
) +

scale_x_continuous(breaks = seq(2010, 2025, 1)) +

labs(
title = "Massachusetts Solar Installations vs SREC Price Range",
subtitle = "Bars = installations | Black line = SREC auction value | Orange band = price range",
x = "Year"
) +

theme_minimal()

```

You can also embed plots, for example:
```{r}
# ================================
# Massachusetts Solar + SREC + SMART Analysis (Corrected)
# ================================

library(dplyr)
library(ggplot2)
library(lubridate)
library(tibble)

# -------------------------------
# 1. SREC PRICE DATA (DOER-BASED)
# -------------------------------
ma_srec <- tibble(
  year = 2010:2025,

  acp_srec1 = c(
    600, 550, 550, 550,
    523, 496, 472, 448,
    426, 404, 384, 365,
    347, 330, 330, 330
  ),

  acp_srec2 = c(
    NA, NA, NA, NA,
    375, 375, 350, 350,
    350, 333, 316, 300,
    285, 271, 257, 244
  ),

  auction_price = c(
    300, 300, 300, 300,
    300, 300, 300,
    285, 271, 257,
    244, 232, 221,
    210, 199, 189
  )
) |>
  mutate(auction_net = auction_price * 0.95)

# -------------------------------
# 2. SMART INCENTIVE (RESIDENTIAL APPROX)
# -------------------------------
smart <- tibble(
  year = 2018:2025,

  # $/kWh adder (approx program averages)
  smart_adder_kwh = c(
    0.10, 0.095, 0.090, 0.080,
    0.070, 0.060, 0.050, 0.045
  )
) |>
  mutate(
    # convert to $/MWh equivalent for comparability
    smart_price_mwh = smart_adder_kwh * 1000
  )

# -------------------------------
# 3. INSTALLATIONS (YEARLY)
# -------------------------------
installs_yearly <- mass_solar_clean |>
  filter(
    facility_type == "Residential (3 or fewer dwelling units per building)",
    year >= 2010
  ) |>
  mutate(year = year(as.Date(date_in_service))) |>
  count(year, name = "num_instal")

# -------------------------------
# 4. MERGE ALL DATA
# -------------------------------
plot_data <- installs_yearly |>
  left_join(ma_srec, by = "year") |>
  left_join(smart, by = "year") |>
  arrange(year)

# -------------------------------
# 5. SCALE FACTOR (SREC + SMART)
# -------------------------------
scale_factor <- max(plot_data$num_instal, na.rm = TRUE) /
  max(
    c(plot_data$auction_net, plot_data$smart_price_mwh),
    na.rm = TRUE
  )

# -------------------------------
# 6. MAIN VISUALIZATION
# -------------------------------
ggplot(plot_data, aes(x = year)) +

  # Installations
  geom_col(aes(y = num_instal),
           fill = "steelblue", alpha = 0.8) +

  # SREC floor price (black line)
  geom_line(
    aes(y = auction_net * scale_factor),
    color = "black",
    linewidth = 1.2
  ) +

  geom_point(
    aes(y = auction_net * scale_factor),
    color = "black"
  ) +

  # SMART incentive (green line, only post-2018 visible)
  geom_line(
    aes(y = smart_price_mwh * scale_factor),
    color = "darkgreen",
    linewidth = 1.2,
    na.rm = TRUE
  ) +

  geom_point(
    aes(y = smart_price_mwh * scale_factor),
    color = "darkgreen",
    na.rm = TRUE
  ) +

  # X-axis readability
  scale_x_continuous(
    breaks = seq(2010, 2025, by = 2)
  ) +

  # Dual axis
  scale_y_continuous(
    name = "Number of Installations",
    sec.axis = sec_axis(~ . / scale_factor,
                        name = "Incentive ($/MWh equivalent)")
  ) +

  labs(
    title = "Massachusetts Solar Installations vs SREC and SMART Incentives",
    subtitle = "Black = SREC payouts | Green = SMART incentives | Bars = installations",
    x = "Year"
  ) +

  theme_minimal()

# -------------------------------
# 7. DOES PRICE DRIVE INSTALLS?
# -------------------------------

analysis_data <- plot_data |>
  mutate(
    srec_price = auction_net,
    srec_price_lag1 = lag(srec_price, 1),

    smart_price = smart_price_mwh,
    smart_price_lag1 = lag(smart_price, 1),

    policy_period = case_when(
      year <= 2014 ~ "SREC-1",
      year <= 2018 ~ "SREC-2",
      TRUE ~ "SMART"
    )
  )

# Correlation (SREC)
cor(analysis_data$num_instal,
    analysis_data$srec_price_lag1,
    use = "complete.obs")

# Regression: SREC effect
model1 <- lm(num_instal ~ srec_price_lag1, data = analysis_data)
summary(model1)

# Regression: policy effect
model2 <- lm(num_instal ~ srec_price_lag1 + policy_period,
             data = analysis_data)
summary(model2)

# Regression: SMART signal
model3 <- lm(num_instal ~ smart_price_lag1, data = analysis_data)
summary(model3)

# -------------------------------
# 8. DIAGNOSTIC PLOT
# -------------------------------
ggplot(analysis_data, aes(x = srec_price_lag1, y = num_instal)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(
    title = "Do SREC Prices Drive Installations?",
    x = "Lagged SREC Price ($/MWh equivalent)",
    y = "Installations"
  ) +
  theme_minimal()
```
```{r}
# ================================
# Massachusetts SREC Price Range (NO installations)
# ================================

library(dplyr)
library(ggplot2)
library(tibble)

# -------------------------------
# 1. SREC price dataset (DOER-based)
# -------------------------------

ma_srec <- tibble(
  year = 2010:2025,

  acp_srec1 = c(
    600, 550, 550, 550,
    523, 496, 472, 448,
    426, 404, 384, 365,
    347, 330, 330, 330
  ),

  acp_srec2 = c(
    NA, NA, NA, NA,
    375, 375, 350, 350,
    350, 333, 316, 300,
    285, 271, 257, 244
  ),

  auction_price = c(
    300, 300, 300, 300,
    300, 300, 300,
    285, 271, 257,
    244, 232, 221,
    210, 199, 189
  )
) |>
  mutate(
    auction_net = auction_price * 0.95,
    acp = ifelse(is.na(acp_srec2), acp_srec1, acp_srec2)
  )

# -------------------------------
# 2. PLOT ONLY SREC PRICE RANGE
# -------------------------------

ggplot(ma_srec, aes(x = year)) +

  # Price band (floor → ceiling)
  geom_ribbon(
    aes(
      ymin = auction_net,
      ymax = acp
    ),
    fill = "orange",
    alpha = 0.4
  ) +

  # Floor price (actual payout)
  geom_line(
    aes(y = auction_net),
    color = "black",
    linewidth = 1.2
  ) +

  geom_point(
    aes(y = auction_net),
    color = "black"
  ) +

  scale_x_continuous(breaks = seq(2010, 2025, 1)) +

  labs(
    title = "Massachusetts SREC Price Evolution (2010–2025)",
    subtitle = "Black = clearing price (floor) | Orange band = policy price range (floor to ACP)",
    x = "Year",
    y = "SREC Price ($/MWh)"
  ) +

  theme_minimal()

```
```{r}
# ================================
# Massachusetts SREC: Correlation + Regression (2010–2018)
# Fully self-contained
# ================================

library(dplyr)
library(lubridate)

# -------------------------------
# 1. SREC price dataset (DOER-based)
# -------------------------------
ma_srec <- tibble(
  year = 2010:2025,

  acp_srec1 = c(
    600, 550, 550, 550,
    523, 496, 472, 448,
    426, 404, 384, 365,
    347, 330, 330, 330
  ),

  acp_srec2 = c(
    NA, NA, NA, NA,
    375, 375, 350, 350,
    350, 333, 316, 300,
    285, 271, 257, 244
  ),

  auction_price = c(
    300, 300, 300, 300,
    300, 300, 300,
    285, 271, 257,
    244, 232, 221,
    210, 199, 189
  )
) |>
  mutate(
    auction_net = auction_price * 0.95,
    acp = ifelse(is.na(acp_srec2), acp_srec1, acp_srec2)
  )

# -------------------------------
# 2. Installations (annual)
# -------------------------------
installs_yearly <- mass_solar_clean |>
  filter(
    facility_type == "Residential (3 or fewer dwelling units per building)"
  ) |>
  mutate(year = year(as.Date(date_in_service))) |>
  filter(year >= 2010, year <= 2018) |>
  count(year, name = "num_instal")

# -------------------------------
# 3. Merge datasets
# -------------------------------
analysis_data <- installs_yearly |>
  left_join(ma_srec, by = "year") |>
  mutate(
    srec_floor = auction_net,
    srec_ceiling = acp,
    srec_mid = (auction_net + acp) / 2
  )

# -------------------------------
# 4. Correlations
# -------------------------------
cor_floor <- cor(analysis_data$num_instal, analysis_data$srec_floor, use = "complete.obs")
cor_ceiling <- cor(analysis_data$num_instal, analysis_data$srec_ceiling, use = "complete.obs")
cor_mid <- cor(analysis_data$num_instal, analysis_data$srec_mid, use = "complete.obs")

cor_floor
cor_ceiling
cor_mid

# -------------------------------
# 5. Linear regressions
# -------------------------------

model_floor <- lm(num_instal ~ srec_floor, data = analysis_data)
summary(model_floor)

model_ceiling <- lm(num_instal ~ srec_ceiling, data = analysis_data)
summary(model_ceiling)

model_range <- lm(num_instal ~ srec_floor + srec_ceiling, data = analysis_data)
summary(model_range)
```
```{r}
library(dplyr)
library(zoo)
library(ggplot2)

# -------------------------------
# 1. Build full dataset (2010–2018 SREC era)
# -------------------------------
ma_srec <- tibble(
  year = 2010:2025,
  acp_srec1 = c(
    600, 550, 550, 550,
    523, 496, 472, 448,
    426, 404, 384, 365,
    347, 330, 330, 330
  ),
  acp_srec2 = c(
    NA, NA, NA, NA,
    375, 375, 350, 350,
    350, 333, 316, 300,
    285, 271, 257, 244
  ),
  auction_price = c(
    300, 300, 300, 300,
    300, 300, 300,
    285, 271, 257,
    244, 232, 221,
    210, 199, 189
  )
) |>
  mutate(
    auction_net = auction_price * 0.95,
    acp = ifelse(is.na(acp_srec2), acp_srec1, acp_srec2)
  )

installs_yearly <- mass_solar_clean |>
  filter(
    facility_type == "Residential (3 or fewer dwelling units per building)"
  ) |>
  mutate(year = lubridate::year(as.Date(date_in_service))) |>
  filter(year >= 2010) |>
  count(year, name = "num_instal")

data <- installs_yearly |>
  left_join(ma_srec, by = "year") |>
  arrange(year)

# -------------------------------
# 2. Rolling correlation (3-year window)
# -------------------------------
data$roll_cor <- zoo::rollapply(
  data = data.frame(data$num_instal, data$auction_net),
  width = 3,
  FUN = function(x) cor(x[,1], x[,2], use = "complete.obs"),
  by.column = FALSE,
  align = "right",
  fill = NA
)

# -------------------------------
# 3. Plot rolling correlation
# -------------------------------
ggplot(data, aes(x = year, y = roll_cor)) +
  geom_line(color = "steelblue", linewidth = 1.2) +
  geom_point(color = "steelblue") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_x_continuous(breaks = seq(2010, 2018, 1)) +
  labs(
    title = "Rolling Correlation: SREC Prices vs Solar Adoption (Massachusetts)",
    subtitle = "3-year moving correlation (auction_net vs installations)",
    x = "Year",
    y = "Correlation"
  ) +
  theme_minimal()
```
```{r}
# ================================
# Massachusetts Solar: Rolling Correlation (2010–2018 only)
# ================================

library(dplyr)
library(lubridate)
library(zoo)
library(ggplot2)
library(tibble)

# -------------------------------
# 1. SREC price dataset
# -------------------------------
ma_srec <- tibble(
  year = 2010:2025,

  acp_srec1 = c(
    600, 550, 550, 550,
    523, 496, 472, 448,
    426, 404, 384, 365,
    347, 330, 330, 330
  ),

  acp_srec2 = c(
    NA, NA, NA, NA,
    375, 375, 350, 350,
    350, 333, 316, 300,
    285, 271, 257, 244
  ),

  auction_price = c(
    300, 300, 300, 300,
    300, 300, 300,
    285, 271, 257,
    244, 232, 221,
    210, 199, 189
  )
) |>
  mutate(
    auction_net = auction_price * 0.95,
    acp = ifelse(is.na(acp_srec2), acp_srec1, acp_srec2)
  )

# -------------------------------
# 2. Installations (STRICT 2010–2018)
# -------------------------------
installs_yearly <- mass_solar_clean |>
  filter(
    facility_type == "Residential (3 or fewer dwelling units per building)"
  ) |>
  mutate(year = year(as.Date(date_in_service))) |>
  filter(year >= 2010, year <= 2018) |>
  count(year, name = "num_instal")

# -------------------------------
# 3. Merge
# -------------------------------
data <- installs_yearly |>
  left_join(ma_srec, by = "year") |>
  arrange(year)

# -------------------------------
# 4. Rolling correlation (3-year window)

roll_mat <- cbind(data$num_instal, data$auction_net)

# -------------------------------
data$roll_cor <- zoo::rollapply(
  roll_mat,
  width = 3,
  FUN = function(x) cor(x[,1], x[,2], use = "complete.obs"),
  by.column = FALSE,
  align = "right",
  fill = NA
)
# -------------------------------
# 5. FORCE correct x-axis range
# -------------------------------
data <- data |>
  filter(year >= 2010, year <= 2018)

# -------------------------------
# 6. Plot
# -------------------------------
ggplot(data, aes(x = year, y = roll_cor)) +
  geom_line(color = "steelblue", linewidth = 1.2) +
  geom_point(color = "steelblue") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_x_continuous(breaks = seq(2010, 2018, 1)) +
  coord_cartesian(xlim = c(2010, 2018)) +
  labs(
    title = "Rolling Correlation: SREC Prices vs Solar Adoption (2010–2018)",
    subtitle = "3-year moving correlation (auction_net vs installations)",
    x = "Year",
    y = "Correlation"
  ) +
  theme_minimal()

```
```{r}
# ================================
# Massachusetts SMART Rates by Policy Era (Approx)
# ================================

library(dplyr)
library(ggplot2)
library(tibble)

# -------------------------------
# 1. SMART dataset (approximation)
# -------------------------------
smart_rates <- tibble(
  year = 2018:2025,

  base_rate = c(
    0.10, 0.095, 0.090, 0.080,
    0.070, 0.060, 0.050, 0.045
  ),

  avg_adder = c(
    0.035, 0.035, 0.035, 0.030,
    0.030, 0.025, 0.025, 0.025
  )
) |>
  mutate(
    total_rate = base_rate + avg_adder,

    smart_phase = case_when(
      year <= 2020 ~ "SMART 1.0",
      year <= 2023 ~ "SMART 2.0",
      TRUE ~ "SMART 3.0"
    )
  )

# -------------------------------
# 2. Plot segmented SMART phases
# -------------------------------
ggplot(smart_rates, aes(x = year, y = total_rate)) +

  geom_line(color = "darkgreen", linewidth = 1.2) +
  geom_point(color = "darkgreen", size = 2) +

  geom_point(aes(color = smart_phase), size = 3) +

  scale_color_manual(
    values = c(
      "SMART 1.0" = "blue",
      "SMART 2.0" = "orange",
      "SMART 3.0" = "red"
    )
  ) +

  scale_x_continuous(breaks = 2018:2025) +

  labs(
    title = "Massachusetts SMART Compensation Rates by Policy Phase",
    subtitle = "Phases reflect declining block structure over time (not official program names)",
    x = "Year",
    y = "Compensation Rate ($/kWh)",
    color = "SMART Phase"
  ) +

  theme_minimal()

```
```{r}
# ================================
# Massachusetts SMART Guideline → Clean Tibble (NO CSV, NO PDF tools)
# Source:
# https://www.mass.gov/doc/capacity-block-base-compensation-rate-and-compensation-rate-adder-guideline-1/download
# ================================

library(dplyr)
library(stringr)
library(tibble)

# -------------------------------
# 1. Download PDF manually or via R
# -------------------------------
url <- "https://www.mass.gov/doc/capacity-block-base-compensation-rate-and-compensation-rate-adder-guideline-1/download"
pdf_file <- "smart_guideline.pdf"

download.file(url, pdf_file, mode = "wb")

# -------------------------------
# 2. Extract raw text using system call (no packages)
# -------------------------------
# Requires poppler installed (pdftotext is standard on most systems)

system(paste("pdftotext", pdf_file, "smart.txt"))

raw_text <- readLines("smart.txt", warn = FALSE)

# -------------------------------
# 3. Clean text
# -------------------------------
clean_lines <- raw_text |>
  str_squish() |>
  discard(~ .x == "")

# -------------------------------
# 4. Extract key rows (utilities + numeric values)
# -------------------------------
smart_rows <- clean_lines |>
  str_subset("Eversource|National Grid|Unitil|\\d+\\.\\d+")

# -------------------------------
# 5. Parse structured fields
# -------------------------------
smart_df <- tibble(raw = smart_rows) |>
  mutate(
    utility = str_extract(raw, "Eversource|National Grid|Unitil"),

    numbers = str_extract_all(raw, "\\d+\\.\\d+|\\d+"),

    text = str_remove_all(raw, "\\d+\\.\\d+|\\d+")
  )

# -------------------------------
# 6. Expand numeric columns safely
# -------------------------------
smart_clean <- smart_df |>
  mutate(
    rate_1 = sapply(numbers, function(x) as.numeric(x[1])),
    rate_2 = sapply(numbers, function(x) as.numeric(x[2])),
    rate_3 = sapply(numbers, function(x) as.numeric(x[3]))
  ) |>
  select(utility, rate_1, rate_2, rate_3, text)

# -------------------------------
# 7. Final tidy dataset
# -------------------------------
smart_clean
```
```{r}


Base_compensation_rates <- read_csv("C:/Users/pschnell/Downloads/Base compensation rates.csv")
View(Base_compensation_rates)

```

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
