---
title: "Peter's Solar Final Draft 5_2"
output: html_document
date: "2026-05-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidycensus)
library(sf)
#library(tmap)
library(tmaptools)
library(DT)
library(ggplot2)
library(tidyverse)
library(janitor)
library(GGally)
library(DT)            # for visualizing data tables 
library(janitor)       # for automatically renaming variables
library(ggmap)         # for mapping points on maps
library(RColorBrewer)  # for color palettes
library(sf)            # for working with spatial data
library(leaflet)       # for highly customizable mapping
library(ggthemes)      # for more themes (including theme_map())
library(gplots)        # for col2hex() function
library(socviz)        # for the US county map
library(usmap)         # for plot_usmap
library(usdata)       # for abbr2state function
library(choroplethr)   # another option to make choropleths
library(choroplethrMaps)
library(maps)
library(rvest)
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r}
#imports Data
mass_solar_messy <- read_csv("C:/Users/pschnell/Downloads/Data Science sust/Solar-PV-Systems-in-MA-as-of-Jan-2025-readable.csv", 
    col_types = cols(`Date In Service` = col_date(format = "%m/%d/%Y"), 
        `Total Cost with Design Fees` = col_number(), 
        `Total Grant` = col_number()))

#cleans names
mass_solar_clean<-mass_solar_messy |>
  janitor::clean_names()|>
   mutate(year=year(date_in_service), month=month(date_in_service),weekday=wday(date_in_service), day=day(date_in_service))

```

## Including Plots

You can also embed plots, for example:

```{r, echo=FALSE}
all_mass_year<-mass_solar_clean|>
  count(year)|>
 mutate(num_instal=n)|>
  select(-"n")|>
  mutate(num_instal_total=cumsum(num_instal))


  
ggplot(all_mass_year, aes(x=year))+
  geom_col(aes(y=num_instal))+
  geom_line(aes(y=num_instal_total))+
 # geom_col(aes(y=count), fill="green")+
  ggtitle("All Solar Registered Instalations in Massachusets ")+
labs(y="Number of Instalations", x="Year", fill="Number covered by grants")
```
```{r}
# install.packages(c("ggplot2", "dplyr", "scales"))
library(ggplot2)
library(dplyr)
library(scales)

# Timeline data
solar_programs <- data.frame(
  program = c(
    "Federal Solar Tax Credit (“Energy Tax Act)",
    "Federal Solar Tax Credit (ITC)",
    "MA Solar Tax Credit",
    "MassCEC Commonwealth Solar Grants",
    "SREC I",
    "SREC II",
    "Mass CEC Solar Loan Program",
    "Mass Smart",
    "Mass CEC EmPower Grants"
  ),

  start = as.Date(c(
    "1978-01-01",
    "2006-01-01",
    "1979-01-01",
    "2008-01-01",
    "2010-01-01",
    "2014-04-25",
    "2015-06-01",
    "2018-11-26",
    "2020-01-01"
  )),

  end = as.Date(c(
    "1985-12-31",
    "2034-12-31",
    "2030-12-31",
    "2015-12-31",
    "2014-04-24",
    "2018-11-26",
    "2020-12-31",
    "2030-12-31",
    "2030-12-31"
  ))
)

# Reverse order for plotting
solar_programs$program <- factor(
  solar_programs$program,
  levels = rev(solar_programs$program)
)

# Every-other-year breaks starting at 1978
year_breaks <- seq(
  as.Date("1978-01-01"),
  as.Date("2034-01-01"),
  by = "2 years"
)

# Plot
ggplot(solar_programs) +

  # Narrow timeline bars
  geom_segment(
    aes(
      x = start,
      xend = end,
      y = program,
      yend = program
    ),
    linewidth = 4,
    color = "blue"
  ) +

  # X-axis formatting
  scale_x_date(
    breaks = year_breaks,
    date_minor_breaks = "1 year",
    labels = label_date("%Y"),
    expand = expansion(mult = c(0.01, 0.01))
  ) +

  labs(
    title = "Massachusetts & Federal Solar Incentives",
    x = "Year",
    y = NULL
  ) +

  theme_minimal(base_size = 12) +

  theme(
    panel.grid.major.x = element_line(
      color = "gray",
      linewidth = 0.45
    ),

    panel.grid.minor.x = element_line(
      color = "gray",
      linewidth = 0.25
    ),

    panel.grid.major.y = element_line(color = "gray"),
    panel.grid.minor.y = element_blank(),

    axis.text.x = element_text(
      angle = 90,
      vjust = 0.5,
      size = 8
    ),

    axis.text.y = element_text(
      face = "bold",
      size = 10
    ),

    plot.title = element_text(
      face = "bold",
      size = 16
    ),

    plot.subtitle = element_text(size = 11)
  )
```
```

```{r}
all_mass_year_type<-mass_solar_clean|>
  group_by(year, facility_type)|>
 summarise(mw=sum(capacity_dc_k_w)/1000)

all_mass_year_mw<-mass_solar_clean|>
  group_by(year)|>
summarise(total_mw=sum(capacity_dc_k_w/1000))|>
 mutate(mw_cum=(cumsum(total_mw)))


all_mass_year_type_mw<-all_mass_year_type|>
  full_join(all_mass_year_mw)

all_mass_year_type_mw

ggplot(all_mass_year_type_mw, aes(x=year))+
  geom_col(aes(y=mw, fill=facility_type))+
  geom_line(aes(y=mw_cum))+
  ggtitle("Capacity of All Registered Solar Instalations in Massachusetts before 2025")+
labs(y="Total Capacity DC (MW)", x="Year", fill="Facility Type")

ggplot(all_mass_year_type_mw, aes(x=year))+
  geom_col(aes(y=mw))+
  geom_line(aes(y=mw_cum))+
  ggtitle("Capacity of All Registered Solar Instalations in Massachusetts before 2025")+
labs(y="Total Capacity DC (MW)", x="Year")

```


```{r, echo=FALSE}
all_mass_year_type<-mass_solar_clean|>
  group_by(year)|>
  count(facility_type)|>
 arrange(desc(n))|>
 mutate(num_instal=n)|>
  select(-"n")

  
ggplot(all_mass_year_type, aes(x=year, y=num_instal, fill=facility_type))+
  geom_col()+
 # geom_col(aes(y=count), fill="green")+
  ggtitle("Number of Registered Solar Instalations in Massachusets ")+
labs(y="Number of Instalations", x="Year", fill="Facility Type")

```



```{r}


all_mass_year_com<-mass_solar_clean|>
  group_by(year,facility_type)|>
 summarise(mw=sum(capacity_dc_k_w)/1000)|>
 arrange(desc(mw))

  
ggplot(all_mass_year_com, aes(x=year, y=mw, fill=facility_type))+
  geom_col()+
 # geom_col(aes(y=count), fill="green")+
  ggtitle("Capacity of All Registered Solar Instalations in Massachusetts before 2025")+
labs(y="Total Capacity DC (MW)", x="Year", fill="Facility Type")


```
tax credit could be added to the federal credit, and had remained steady since it was passed, and despite federal funding cuts. Sources: (mass solar) (Read more about Massachusetts tax credit here)


```{r}
all_mass_year_type<-mass_solar_clean|>
  filter(facility_type=="Community Solar")|>
  group_by(year)|>
 summarise(mw=sum(capacity_dc_k_w)/1000)|>
 arrange(desc(mw))

  
ggplot(all_mass_year_type, aes(x=year, y=mw))+
  geom_col()+
 # geom_col(aes(y=count), fill="green")+
  ggtitle("Capacity of registered Community Solar Instalations Over Time")+
labs(y="Total Capacity DC (MW)", x="Year", fill="Facility Type")
```

```{r}
all_mass_year_com1<-mass_solar_clean|>
   filter(facility_type=="Community Solar")|>
  group_by(year)|>
  count()|>
 arrange(desc(n))|>
 mutate(num_instal=n)|>
  select(-"n")

  
ggplot(all_mass_year_com1, aes(x=year, y=num_instal))+
  geom_col()+
 # geom_col(aes(y=count), fill="green")+
  ggtitle("Number of Registered Community Solar Instalations in Massachusets ")+
labs(y="Number of Instalations", x="Year", fill="Facility Type")
```
```{r}
mass_com_solar_ave<-mass_solar_clean|>
  filter(facility_type=="Residential (3 or fewer dwelling units per building)")|>
  group_by(year)|>
  summarise(mean=mean(capacity_dc_k_w))



ggplot(mass_com_solar_ave, aes(x=year, y=mean))+
  geom_line()+
  ggtitle("Average Capacity of Registered Community Solar Instalations in Massachusets ")+
 labs(y="Capacity of Instalations (MW)", x="Year", )

```
```{r}
mass_com_solar_ave<-mass_solar_clean|>
  filter(facility_type=="Community Solar")|>
  group_by(year)|>
  summarise(mean=mean(capacity_dc_k_w))

ggplot(mass_com_solar_ave, aes(x=year, y=mean))+
  geom_line()+
  ggtitle("Average Capacity of Registered Community Solar Instalations in Massachusets ")+
 labs(y="Capacity of Instalations (MW)", x="Year", )
```


```{r}
all_mass_year_mw<-mass_solar_clean|>
  group_by(year)|>
summarise(total_mw=sum(capacity_dc_k_w/1000))|>
 mutate(mw=(cumsum(total_mw)))
  
ggplot(all_mass_year_mw)+
  geom_line(aes(x=year, y=mw))+
 # geom_col(aes(y=count), fill="green")+
  ggtitle("Capacity of All Registered Solar Instalations in Massachusetts before 2025")+
labs(y="Total Capacity DC (MW)", x="Year")

```
```{r}
all_mass_year_type<-mass_solar_clean|>
  group_by(year, facility_type)|>
 summarise(mw=sum(capacity_dc_k_w)/1000)

all_mass_year_mw<-mass_solar_clean|>
  group_by(year)|>
summarise(total_mw=sum(capacity_dc_k_w/1000))|>
 mutate(mw_cum=(cumsum(total_mw)))


all_mass_year_type_mw<-all_mass_year_type|>
  full_join(all_mass_year_mw)

all_mass_year_type_mw

ggplot(all_mass_year_type_mw, aes(x=year))+
  geom_col(aes(y=mw, fill=facility_type))+
  geom_line(aes(y=mw_cum))+
  ggtitle("Capacity of All Registered Solar Instalations in Massachusetts before 2025")+
labs(y="Total Capacity DC (MW)", x="Year", fill="Facility Type")

ggplot(all_mass_year_type_mw, aes(x=year))+
  geom_col(aes(y=mw))+
  geom_line(aes(y=mw_cum))+
  ggtitle("Capacity of All Registered Solar Instalations in Massachusetts before 2025")+
labs(y="Total Capacity DC (MW)", x="Year")

```
```{r}

all_mass<- mass_solar_clean|>
  filter(facility_type=="Residential (3 or fewer dwelling units per building)") |>
  mutate(has_grant= total_grant>0)|>
  arrange(desc(total_grant))

all_mass_year<-all_mass|>
  count(year)|>
 mutate(num_instal=n)|>
  select(-"n")

grants_all<-all_mass|>
  group_by(year)|>
  summarise(count=sum(has_grant, na.rm=TRUE))
  
all_grant_year<-all_mass_year|>
  left_join(grants_all)
  
ggplot(all_grant_year, aes(x=year))+
  geom_col(aes(y=num_instal))+
  geom_col(aes(y=count), fill="green")+
  ggtitle("all single family solar instalations over time, with grant recieved ")+
labs(y="Number of Instalations", x="Year", fill="Number covered by grants")


all_mass_month<-all_mass|>
  mutate(year_month=sub("^([^-]*-[^-]*)-.*", "\\1", date_in_service))


all_mass_month1<-all_mass_month|>
  count(year_month)|>
 mutate(num_instal=n)|>
  select(-"n")

grants_all<-all_mass_month|>
  group_by(year_month)|>
  summarise(count=sum(has_grant, na.rm=TRUE))
  
all_grant_month1<-all_mass_month1|>
  left_join(grants_all)

  
library (plotly)
all<-ggplot(all_grant_month1, aes(x=year_month))+
  geom_col(aes(y=num_instal))+
  geom_col(aes(y=count), fill="green")+
  ggtitle("all single family solar instalations over time, with grant recieved ")+
labs(y="Number of Instalations", x="month", fill="Number covered by grants")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
ggplotly(all)

```
```{r}
all_mass<- mass_solar_clean|>
  filter(facility_type=="Residential (3 or fewer dwelling units per building)") |>
  mutate(has_grant= total_grant>0)|>
  arrange(desc(total_grant))

all_mass_month<-all_mass|>
  mutate(year_month=sub("^([^-]*-[^-]*)-.*", "\\1", date_in_service))


all_mass_month1<-all_mass_month|>
  count(year_month)|>
 mutate(num_instal=n)|>
  select(-"n")


all_mass_month2<-all_mass_month1|>
  mutate(Solar_insentive_type = case_when(
   between(year_month, "2010-01", "2014-04")~"SREC-1", 
   between(year_month, "2014-05", "2018-11")~"SREC-2",
    between(year_month, "2018-12", "2020-06")~"SMART 1.0",
      between(year_month, "2020-07", "2025-08")~"SMART 2.0",
    between(year_month, "2025-09", "2026-09")~"SMART 3.0"))
  
  
  
library (plotly)
all<-ggplot(all_mass_month2, aes(x=year_month))+
  geom_col(aes(y=num_instal, fill= Solar_insentive_type))+
  ggtitle("all single family solar instalations over time, with insentive period ")+
labs(y="Number of Instalations", x="month", fill="Number covered by grants")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
ggplotly(all)


```
```{r}
all_mass<- mass_solar_clean|>
  filter(facility_type=="Residential (3 or fewer dwelling units per building)") |>
  mutate(has_grant= total_grant>0)|>
  filter(year<="2008")|>
  arrange(desc(total_grant))

all_mass_month<-all_mass|>
  mutate(year_month=sub("^([^-]*-[^-]*)-.*", "\\1", date_in_service))


all_mass_month1<-all_mass_month|>
  count(year_month)|>
 mutate(num_instal=n)|>
  select(-"n")


all_mass_month2<-all_mass_month1|>
  mutate(Solar_insentive_type = case_when(
   between(year_month, "2010-01", "2014-04")~"SREC-1", 
   between(year_month, "2014-05", "2018-11")~"SREC-2",
    between(year_month, "2018-12", "2020-06")~"SMART 1.0",
      between(year_month, "2020-07", "2025-08")~"SMART 2.0",
    between(year_month, "2025-09", "2026-09")~"SMART 3.0"))
  
  
  
#library (plotly)
#all<-
  ggplot(all_mass_month2, aes(x=year_month))+
  geom_col(aes(y=num_instal, fill= Solar_insentive_type))+
  ggtitle("all single family solar instalations over time, with insentive period ")+
labs(y="Number of Instalations", x="month", fill="Number covered by grants")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
#ggplotly(all)


```
```{r}
#from chat GPET

ma_srec <- tibble(
  year = 2010:2025,

  acp_srec1 = c(
    600, 550, 550, 550,
    523, 496, 472, 448,
    426, 404, 384, 365,
    347, 330, 330, 330
  ),

  acp_srec2 = c(
    NA, NA, NA, NA,
    375, 375, 350, 350,
    350, 333, 316, 300,
    285, 271, 257, 244
  ),

  auction_price = c(
    300, 300, 300, 300,
    300, 300, 300,
    285, 271, 257,
    244, 232, 221,
    210, 199, 189
  )
) |>
  mutate(
    auction_net = auction_price * 0.95,
    acp = ifelse(is.na(acp_srec2), acp_srec1, acp_srec2)
  )


ggplot(ma_srec, aes(x = year)) +

  # Price band (floor → ceiling)
  geom_ribbon(
    aes(
      ymin = auction_net,
      ymax = acp
    ),
    fill = "orange",
    alpha = 0.4
  ) +

  # Floor price (actual payout)
  geom_line(
    aes(y = auction_net),
    color = "black",
    linewidth = 1.2
  ) +

  geom_point(
    aes(y = auction_net),
    color = "black"
  ) +

  scale_x_continuous(breaks = seq(2010, 2025, 1)) +

  labs(
    title = "Massachusetts SREC Price Evolution (2010–2025)",
    subtitle = "Black = clearing price (floor) | Orange band = policy price range (floor to ACP)",
    x = "Year",
    y = "SREC Price ($/MWh)"
  ) +

  theme_clean()
```


```{r}
# install.packages("ggplot2")

library(ggplot2)

# SMART incentive values
values <- c(
  0.31126, 0.29881, 0.28686, 0.27538,
  0.26437, 0.25379, 0.24364, 0.23390,
  0.22454, 0.21556, 0.20694, 0.19866,
  0.19071, 0.18308, 0.17576, 0.16873
)

# Create dataframe
smart_df <- data.frame(
  Block = 1:16,
  Incentive = values
)

# Plot
ggplot(smart_df, aes(x = Block, y = Incentive)) +

  geom_line(
    color = "steelblue",
    linewidth = 1.2
  ) +

  geom_point(
    color = "darkblue",
    size = 2.5
  ) +

  geom_text(
    aes(label = round(Incentive, 3)),
    vjust = -0.8,
    size = 3
  ) +

  scale_x_continuous(
    breaks = 1:16
  ) +

  labs(
    title = "Mass SMART Incentive Decline by Block",
    subtitle = "≤ 25 kW AC — 200% Compensation Rate — 10-Year Term",
    x = "Capacity Block",
    y = "Incentive Rate ($/kWh)"
  ) +

  theme_minimal(base_size = 13) +

  theme(
    panel.grid.minor = element_line(color = "grey92"),
    panel.grid.major = element_line(color = "grey80"),

    plot.title = element_text(
      face = "bold",
      size = 16
    ),

    axis.text = element_text(size = 10)
  )
```

```{r}
	Less than or equal to 25 kW AC	200%	10-year	$0.31126	$0.29881	$0.28686	$0.27538	$0.26437	$0.25379	$0.24364	$0.23390	$0.22454	$0.21556	$0.20694	$0.19866	$0.19071	$0.18308	$0.17576	$0.16873

```

```{r}

all_mass_month<-mass_solar_clean|>
   filter(year>="2008")|>
  mutate(year_month=sub("^([^-]*-[^-]*)-.*", "\\1", date_in_service))
 

all_mass_month1<-all_mass_month|>
  count(year_month, facility_type)|>
 mutate(num_instal=n)|>
  select(-"n")


all_mass_month1<-all_mass_month|>
  group_by(year_month,facility_type)|>
 summarise(mw=sum(capacity_dc_k_w)/1000)|>
 arrange(desc(mw))


all_mass_month2<-all_mass_month1|>
  mutate(Solar_insentive_type = case_when(
   between(year_month, "2010-01", "2014-04")~"SREC-1", 
   between(year_month, "2014-05", "2018-11")~"SREC-2",
    between(year_month, "2018-12", "2020-06")~"SMART 1.0",
      between(year_month, "2020-07", "2025-08")~"SMART 2.0",
    between(year_month, "2025-09", "2026-09")~"SMART 3.0"))
  
  
  
library (plotly)
all<-ggplot(all_mass_month2, aes(x=year_month))+
  geom_col(aes(y=mw, color= Solar_insentive_type, fill=facility_type ))+
  ggtitle("Capacity of solar instalations over time, with insentive period ")+
labs(y="Number of Instalations", x="month", fill="Number covered by grants")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
ggplotly(all)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
